Stack Safety is a Security Property

### 2021-04-11 tex/main.tex
Indentation fix

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
  to leave it out of the title but hit it strongly in the abstract.}%
What is stack safety? The phrase is associated with a variety of compiler,
run-time, and hardware mechanisms for guarding against corruption of stack
memory.  But these mechanisms mostly lack rigorous specifications, appealing
either to intuitive expectations of how good stack-using programs should
behave or to specific examples of bad behavior.

We propose a formal criterion for stack safety, phrased in terms of
language-based security: a combination of an integrity property (``each
caller's stack frame is invariant by the callee'') and a confidentiality
property (``the callee's behavior is insensitive to differences in the
caller's state''). These are defined in terms of a control flow model
in which callees are meaningfully distinct from their callers due to being
bracketed by calls and returns. We further apply a straightforward notion of
``well-bracketed control flow'' that ensures that callers' control flow
can continue predictably after the return.

  \bcp{I am still not clear
  how well-bracketedness fits into the story.  (For me, it is the bracketing
  that {\em defines} where calls and returns happen, so there is something
  tautological about saying that `` well-bracketedness condition ensures
  that callers and callees remain meaningfully distinct''!)}
  \sna{Attempted to resolve this, it may need some work.}

Our criterion is inspired by hardware tagging mechanisms for stack
protection, in particular the stack-safety ``micro-policies''
of~\citet{DBLP:conf/sp/RoesslerD18}, but its core reflects general
principles of stack safety\bcp{I wish we did not have to be so apologetic
  about being tied to micro-policies (and it is still not clear to me that
  we have to be... do we have a clear statement someplace of why our
  definition would not apply to CFI, for example? This seems like a critical
  discussion...)}. Finally, we show that micro-policies do indeed enforce
our properties using property-based randomized testing. An eager micropolicy
that catches violations immediately enforces our stronger property, and we
define a weaker, more extensional form of the property that is enforced
by a ``lazy'' micropolicy -- one that, for performance reasons, may permit
local violations as long as they do not influence future behavior.

@@

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, leading to consequences
from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
The protections offered by such mechanisms are generally described in terms
of a set of concrete examples of attacks that they can prevent, leaving a
more general characterization to the reader's intuition.
Moreover, the details of these mechanisms can be quite intricate
and often involve trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

So how should we formally define stack safety? Definitions in the literature are commonly
informal and proceed by enumerating the specific bad things that a given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.\bcp{Most of that sentence should be merged
  into the similar sentence above.  Doesn't belong here.}
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of which we are aware is in
recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (or don't return at all).
\citet{Skorstengaard+19} formalize these properties by defining an idealized machine
in which they visibly hold by construction; an enforcement mechanism (in their case, hardware
capabilities)
can then be shown to provide stack safety by giving a fully abstract embedding function
from the idealized machine to an ordinary machine equipped with the enforcement.
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

\bcp{If we end up adopting my recent proposal for defining subtraces and
  separating WBCF, I think we can be a bit stronger here: We define stack
  safety in IFC terminology based on the same intuitions as their LSE, but
  we consider (and define) WBCF to be a separate thing---desirable, but
  distinct from safety.  (If we make this argument, we'll need to show an
  example of a program that is stack safe but doesn't have WBCF.)}
In this work, we take the (informal) definitions of WBCF and LSE as a starting
point, but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}. We define
trace-based properties of \emph{stepwise stack integrity} and
\emph{lockstep stack confidentiality}.
In applying these security concepts, we treat each caller-callee pair as
mutually distrustful, with the caller needing protection from the callee. \apt{Need more explanation of
  ``nesting'' and our novelty factor here? Also about domains??}
This differs from common instances in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.
\sna{Thinking about how to express domains, or whether to. I'm not sure it's
  an interesting detail at this level. Also not sure how strongly to talk about this novelty.}
{\em Integrity} says that a caller's local data is protected from modification by its callee,
and its {\em stepwise} interpretation formalizes this very naturally as a caller's local data
never being modified during a call. {\em Confidentiality} says that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of subtraces for each call, nested within other calls. The term
{\em lockstep} describes the behavior of these pairs: they step together, and any
deviation is recognized as a violation of confidentiality.
Together, these properties capture the rough intuition of LSE, but are not identical --
confidentiality contains temporal protections that LSE does not address.

Since our stack safety occupies the same conceptual space as LSE, we should consider how it
relates to WBCF. First, we observe that some degree of bracketed control flow is implicit
in any notion stack safety, simply because calls and returns must be recognizeable; callers
and callees, distinct. We treat control flow that creates an ambiguity between caller and
callee as keeping the callee in control, and therefore such behavior can never give access
to the caller's data. An example would be if the callee attempts to ``return'' to the
wrong location, hoping to execute code on behalf of the caller: it may make the jump
without violating stack safety, but it gains no privileges on the caller's data.
Alternatively a callee might return past its immediate caller, to somewhere deeper in the
stack, and stack safety does not inherently rule this out. We then define WBCF as an additional
restriction, describing the standard control flow principle that a return pops exactly one frame
from the stack.
\sna{We got side-tracked a bit from talk of entry integrity, and Andrew made some good points
  about it, so let's not rule that out just yet. But I focused on returns because there's a
  clear story already.}

Part of the challenge in fully characterizing stack safety is that at a low level,
without built-in stacks, we are at the mercy of architectural choices that
do not relate to the abstract notion of stack safety. We abstract over these choices
as much as possible, so that our model can be applied to a wide range of architectures
and enforcement mechanisms. But a second challenge is the diversity of enforcement
mechanisms and their need for efficiency. In the name of efficiency, enforcement
mechanisms may weaken their protections, arguing that the protections they do provide
still capture stack safety in a meaningful sense. This calls for an {\em extensional}
characterization of stack safety, one that describes the externally visible behavior
of a stack safe system. Our stepwise and lockstep properties are {\em intensional},
focusing on the internal behavior.

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
Intuitively, memory cells within each stack frame are tagged with the
identity of the function activation
that owns the frame, and access to those locations is then gated
according to the activation that is currently executing. Roessler and Dehon
explore alternative versions of this basic idea, with different ways of identifying
functions and deciding exactly which accesses are allowed, to enable different ways
of maintaining the connection between an activation identity and the memory it owns.
These vary in their efficiency and in the spatial and temporal precision of their
security guarantees.
Their \emph{Depth Isolation} policy requires a strict mapping from a function
to its frame from the start of the call to its end, and permits the function to write
only into that frame.
This seems simple and intuitively correct, but it can be expensive to implement the strict
mapping. Conversely, their \emph{Lazy Tagging and Clearing} policy improves
performance by assigning locations to the active function when they are written, which
permits  ``illicit'' writes to inaccessible stack frames so long as the
written locations are not subsequently read by their ``real'' owners. The justification
for this form of laziness is that any genuinely dangerous writes will ultimately be
detected before they can affect program behavior.

To address this, we develop an \emph{observational integrity} property that
is supports enforcement mechanisms that detect violations between the point when
they occur and the point when they make an observable impact. Intuitively this property
can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
integrity implies observational integrity.
\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
  established that this was important.}
We additionally sketch an associated {\em observational confidentiality} property
makes confidentiality extensional in the same way, though we do not know of a real
enforcement mechanism that weakens confidentiality in this way. Together,
these form an {\em observational stack safety} that is highly extensional.
We therefore argue that, despite its somewhat greater complexity,
it may be a better minimal specification for
stack safety enforcement mechanisms in general.

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
them. Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If the testing can differentiate correct implementations from buggy ones,
it is an argument in favor of their correctness.
%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}
Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy does not enforce stepwise integrity.
Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce observable confidentiality either. %, as we show in
%testing~\apt{really??},
But formalizing the ideal property allows us to see how it falls short -- a specific
violation of the temporal aspect of confidentiality.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
observational integrity, at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
  pretty much done}
\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
  need a caveat about how it's harder to test?}

Our initial description of these properties assumes a single simple stack with no sharing between callers and
callees: all parameters and return values are passed in registers.  However,
we can refine it to allow passing of scalar stack data, granting a callee
access to the caller's data. (We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model.\bcp{This is all a bit mysterious} While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. A tag-based approach can easily
compose multiple such policies.)
\rb{So no pointer passing? Pass by reference is now discussed at the end of
  Section 3, however it looks like things get tricky rather quickly.}
\sna{The trick is that if a programmer is passing pointers, we can't presume their
  intent -- how long should we permit memory to be shared? But if the compiler is passing
  pointers, as in pass by reference, we have a clear sense of where sharing should stop.}

@@

While our property definitions were developed with micro-policy enforcement
in mind, they agnostic about many implementation details and should be
applicable to other systems with minor adjustment.\apt{move this up?} \bcp{I
  am still hoping to get a clearer understanding for myself of just what is
  micropolicy-specific about our definitions...}

@@

In summary, we offer the following contributions: \bcp{Section numbers are
  wonky---why just 7.1, where is 8, etc.}
\begin{itemize}
\item
  We adapt formulations of integrity and confidentiality from
  language-based security
  to formalize a parameterized {\em stepwise stack-safety property} and argue that
  it captures the intuitive concepts of local stack encapsulation, well-bracketed control
  flow, and temporal stack safety~(\cref{sec:lse-and-wbcf}).
\item
  We use a property-based randomized testing framework to support the claim
  that this stepwise property is actually enforced by the Depth Isolation policy
  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:conservative}).
\item
  We formalize a weaker but more extensional \emph{observational} variant of the
  stack safety property and argue that it accurately characterizes
  the Lazy Tagging and Clearing policy of~\citet{DBLP:conf/sp/RoesslerD18}
  enhanced with unique tags on each activation~(\cref{sec:lazy}).
\item
  We show how our models can be instantiated and extended to handle passing
  arguments on the stack~(\cref{sec:??}) and simple
  coroutines~(\cref{sec:??}).\bcp{``instantiated and extended'' is awkward.
  Maybe ``We show how our models can be instantiated to handle passing
  arguments on the stack~(\cref{sec:??}) and extended with simple
  coroutines~(\cref{sec:??}).''  If that's what we mean.}
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micro-policy enforcement
%  mechanism~(\Cref{sec:micro-policy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work, and
\cref{sec:future} sketches directions for future work.

We begin with a description of our assumptions and threat model (\cref{sec:threat}),
then show how our properties apply to examples of stack unsafety 
(\cref{sec:running-example}). Then, after some technical preliminaries
(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}, \cref{sec:lazy})
and describe testing (\cref{sec:testing}).

@@

\bcp{Again, I think we need to put ``Assumptions'' (i.e., our basic picture
  of the world) first, as context.  E.g., none of the threat model
  discussion makes sense until the reader understands that we {\em define}
  function calls and returns by means of annotations (and that we MUST do
  something like that because we are talking about machine code).}

@@

Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the machine that instantiates it.  We assume a register
machine with a program counter \(\PCname\) and stack pointer \(\rsp\); other
details of the machine's ISA are unimportant. We assume that we know ahead of time
the layout of memory, including code and which addresses are within
the bounds of the stack.

Our stack-safety concepts depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code (the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes).
%
We assume we are given a machine language program annotated with the locations of
instructions representing calls, returns, and function entry points. Although this
information would typically come from a compiler together with the machine code,
we do not assume anything about their provenance.

We further parameterize over the structure of the stack, particularly the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention} and model it as a relation between machine states
and addresses. Then from any call, those the sealing convention relates to the
state will be protected until the return.

@@

Intuititively, stack safety means that a caller's stack frame is protected
from a confused or malicious callee's attempts to read or write its private data.
The attacker model therefore treats every caller as a defender and its callee as
an attacker, meaning that with nested calls, a single function may play both roles
at varying points.

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
%  to leave it out of the title but hit it strongly in the abstract.}%
What is stack safety? The phrase is associated with a variety of compiler,
run-time, and hardware mechanisms for guarding against corruption of stack
memory.  But these mechanisms typically lack rigorous specifications;
they rely instead on informal descriptions of
the safe stack manipulation behaviors that they allow,
or specific examples of the bad behaviors that they prevent.

We propose a formal criterion \apt{criteria?} for stack safety, phrased in terms of
language-based security: a combination of an integrity property (``each
caller's stack frame is held invariant by the callee'') and a confidentiality
property (``the callee's behavior is insensitive to differences in the
caller's state''). These properties are sensible even in the presence of
abnormal control flow.\apt{Is this important enough to mention here?}
We further apply\apt{what does this mean? we didn't
  invent it} a straightforward notion of
``well-bracketed control flow'' that restricts control flow to the
typical pattern of matched calls and returns.

We show that our properties describe the behavior of the stack-safety
``micro-policies'' of~\citet{DBLP:conf/sp/RoesslerD18}
using property-based randomized testing. An eager micro-policy
that catches violations immediately enforces our primary ``stepwise''
properties, and we define weaker, more extensional ``observational''
properties that are enforced even by a ``lazy'' micro-policy -- one
that, for performance reasons, may catch violations after they occur
but before they become visible.

We argue \apt{do we?} that our properties are suited to a range of other enforcement
mechanisms. The strict stepwise properties can describe mechanisms
that catch errors immediately, as is typical. While the observational
properties are closely linked to lazy micro-policies they are weak enough
to be a lower bound that captures stack safety for any reasonable enforcement mechanism.

\apt{This is a good story, but does it describe the paper we are able to produce this week?
  We were rejected before exactly because we claimed to have
  a general criterion, but did not in fact demonstrate its generality (which is not
  obvious). What has changed?}

@@

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, leading to consequences
that range from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).}
\fi
%
The protections offered by such mechanisms are generally described in terms
of a set of concrete examples of attacks that they can prevent, such as corruption
of return addresses, buffer overflows, or use of uninitialized variables. This leaves the
task of formulating a more general characterization to the reader's intuition.
Moreover, the details of these mechanisms can be quite intricate
and often involve trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of which we are aware is in
recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (or don't return at all).
\citet{Skorstengaard+19} formalize these properties by defining an idealized machine
in which they visibly hold by construction; an enforcement mechanism (in their case, hardware
capabilities)
can then be shown to provide stack safety by giving a fully abstract embedding function
from the idealized machine to an ordinary machine equipped with the enforcement.
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In this work, we take the (informal) definitions of WBCF\apt{or maybe not?}  and LSE as a starting
point, but choose a different approach to formalizing them, based on the
technical framework of language-based security~\citep{??}. We define
trace-based properties of \emph{stepwise stack integrity} and
\emph{lockstep stack confidentiality}.\apt{Could we just say ``stepwise'' for both?}
In applying these security concepts, we treat each caller as needing protection
from its callee, even as that callee may need protection from its callee in turn.
This differs from common practice in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.
\sna{How novel is this nesting?} \apt{Lost the reasoning principle that was here?}
{\em Integrity} says that a caller's local data is protected from modification by its callee,
and its {\em stepwise} interpretation formalizes this very naturally: a caller's local data
is never modified during a call. {\em Confidentiality} says that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of subtraces for each call, nested within other calls. The term
{\em lockstep} describes the behavior of these pairs: they step together, and any
deviation is recognized as a violation of confidentiality.
Together, these properties capture the rough intuition of LSE, but are somewhat stronger:
confidentiality contains temporal protections that LSE does not address.\apt{Maybe want to
  say a little more about this here, since it is a novelty wrt/ existing work?}

Since our stack safety occupies the same conceptual space as LSE, we should consider how it
relates to WBCF.\apt{Or maybe not. This disucssion seems too detailed for this spot in the intro.
  While it is good to note that our integrity/confidentiality properties make sense without
  WBCF-or-similar, LSE is also described independently from WBCF, so there's no novelty in that.}
First, we observe that some degree of bracketed control flow is implicit
in any notion of stack safety, simply because calls and returns must be recognizeable so
that callers and callees are distinguished. We treat control flow that creates an ambiguity between caller and
callee\apt{this is obscure}  as keeping the callee in control, and therefore such behavior can never give access
to the caller's data. An example would be if the callee attempts to ``return'' to the
wrong location, hoping to execute code on behalf of the caller: it may make the jump
without violating stack safety, but it gains no privileges on the caller's data.
Alternatively a callee might return past its immediate caller, to somewhere deeper in the
stack, and stack safety does not inherently rule this out. We therefore define WBCF as an additional
restriction \apt{on what? in what way is it formalized differently than by Skorstengaard?},
describing the standard control flow principle that a return pops exactly one frame
from the stack.
\sna{We got side-tracked a bit from talk of entry integrity, and Andrew made some good points
  about it, so let's not rule that out just yet. But I focused on returns because there's a
  clear story already.}

Part of the challenge in fully\apt{omit?} characterizing stack safety is that in a low-level machine
without built-in stacks, we are at the mercy of architectural
choices that do not relate to the abstract notion of stack safety\apt{a bit vague. what kind of choices?}.
We abstract over these choices
as much as possible, so that our model can be applied to a wide range of architectures
and enforcement mechanisms.

\apt{Intra-paragraph flow seemed awkward.}
A second challenge is
% the diversity of enforcement mechanisms APT:already implied above
enforcement mechanisms may weaken their protections that in the name of efficiency,
while arguing that they still capture stack safety in a meaningful sense.
\apt{Remainder of this paragraph might be better \emph{after} the R+DH para below.}
This calls \apt{not
  immediately obvious why. is this true for anything other than laziness?}
for an {\em extensional}
characterization of stack safety that describes the externally observable behavior
of a stack-safe system. In contract, our stepwise and lockstep properties are {\em intensional},
focusing on the internal behavior.

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
Memory cells within each stack frame are tagged with the
identity of the function activation
that owns the frame, and access to those locations is then gated
according to the activation that is currently executing. Roessler and DeHon
explore alternative versions of this basic idea, with different ways of identifying
functions and deciding exactly which accesses are allowed.
% APT: seem repetitive:
%, to enable different ways
%of maintaining the connection between an activation identity and the memory it owns.
These vary in their efficiency and in the spatial and temporal precision of their
security guarantees.
The \emph{Depth Isolation} policy requires a strict mapping from a function
to its frame from the start of the call to its end\apt{unclear. why not simply say
  what depth is?}, and permits the function to write
only into that frame.
This seems simple and intuitively correct, but it can be expensive to implement the strict
mapping. Conversely, the \emph{Lazy Tagging and Clearing} policy improves
performance by assigning locations to the active function when they are written, which
permits  ``illicit'' writes to inaccessible stack frames so long as the
written locations are not subsequently read by their ``real'' owners. The intuitive justification
for this form of laziness is that any genuinely dangerous writes will ultimately be
detected before they can affect program behavior.

To formalize this intuition, we develop an \emph{observational integrity} property which
characterizes enforcement mechanisms that detect violations between the point when
they occur and the point when they make an observable impact. Intuitively this property
can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
integrity implies observational integrity.
\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
  established that this was important.\apt{medium importance. it is obvious that
    observational integerity allows behaviors that stepwise does not, but
    maybe not obvious that the converse is not true too.}}
We additionally sketch an associated {\em observational confidentiality} property
makes confidentiality extensional in the same way, although we do not know of a real
enforcement mechanism that weakens confidentiality in this way.\apt{why bother mentioning it then?}
Together, these form an {\em observational stack safety} that is highly extensional.
We therefore argue that, despite its somewhat greater complexity,
it may be a better minimal specification for
stack safety enforcement mechanisms in general.

To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
them. Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If the testing can differentiate correct implementations from buggy ones,
it is an argument in favor of their correctness.
%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}
Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy does not enforce stepwise integrity.
Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce observable confidentiality either. %, as we show in
%testing~\apt{really??},
But formalizing the ideal property allows us to see how it falls short -- a specific
violation of the temporal aspect of confidentiality.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
observational integrity, at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
  pretty much done}
\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
  need a caveat about how it's harder to test?}\apt{As per email discussion,
  maybe we could have a property variant that \emph{does} match depth-based Lazy Tagging and Clearing,
  by loosening the temporal safety properties?}

Our initial description of these properties assumes a single simple stack with no sharing between callers and
callees: all parameters and return values are passed in registers.  However,
we can refine it to allow passing of scalar stack data, granting a callee
access to the caller's data. (We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model.\bcp{This is all a bit mysterious} While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. A tag-based approach can easily
compose multiple such policies.)
\rb{So no pointer passing? Pass by reference is now discussed at the end of
  Section 3, however it looks like things get tricky rather quickly.}
\sna{The trick is that if a programmer is passing pointers, we can't presume their
  intent -- how long should we permit memory to be shared? But if the compiler is passing
  pointers, as in pass by reference, we have a clear sense of where sharing should stop.
  \apt{N.B.Doing call-by-reference properly requires that the passed variable is accessible
    by all descendents, not just immediate callee, which is doable with depth tags but not
    obviously with per-activation tags }}

@@

@@

In summary, we offer the following contributions: \bcp{Section numbers are
  wonky---why just 7.1, where is 8, etc.}
\begin{itemize}
\item
  We adapt formulations of integrity and confidentiality from
  language-based security
  to formalize a parameterized {\em stepwise stack-safety property} \apt{lockstep?} and argue that
  it captures the intuitive concepts of local stack encapsulation, well-bracketed control
  flow, and temporal stack safety~(\cref{sec:lse-and-wbcf}).
\item
  We use a property-based randomized testing framework to support the claim
  that this stepwise property is actually enforced by the Depth Isolation policy
  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:conservative}).
\item
  We formalize a weaker but more extensional \emph{observational} variant of the
  stack safety property and argue that it accurately characterizes
  the Lazy Tagging and Clearing policy of~\citet{DBLP:conf/sp/RoesslerD18}
  enhanced with unique tags on each activation~(\cref{sec:lazy}).
\item
  We extend our model so that it can be instantiated in systems that pass arguments
  on the stack~(\cref{sec:??}) and in those with simple coroutines~(\cref{sec:??}).
\end{itemize}

We begin with a description of our assumptions and threat model (\cref{sec:threat}),
then show how our properties apply to examples of stack unsafety 
(\cref{sec:running-example}). Then, after some technical preliminaries
(\cref{sec:prelim}), we formalize the properties (\cref{sec:eager}, \cref{sec:lazy})
and describe testing (\cref{sec:testing}).
\Cref{sec:relwork} discusses related work, and
\cref{sec:future} sketches directions for future work.

@@

@@

Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the concrete machine that instantiates it.  We assume a register
machine with a program counter \(\PCname\) and stack pointer \(\rsp\)\apt{do we really
  need these except for WBCF?}; other
details of the machine's ISA are unimportant. We assume that we know ahead of time
the layout of memory, including where code lives and which addresses are within
the bounds of the stack.

Our stack-safety concepts depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program annotated with the locations of
instructions representing calls, returns, and function entry points. Although these
annotations would typically come from a compiler together with the machine code,
we do not assume anything about their provenance.

We further parameterize over the structure of the stack, particularly the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention} and model it as a relation between machine states
and addresses. From any call, those addresses that the sealing convention relates to the
state\apt{rephrased, but still obscure} will be protected until the return.

@@

Intuitively, stack safety means that a caller's stack frame is protected
from a confused or malicious callee's attempts to read or write its private data.
The attacker model therefore treats every caller as a defender and its callee as
an attacker, meaning that with nested calls, a single function may play both roles
at varying points.
